A Partitioning Algorithm for Maximum Common Subgraph Problems\thanks{This work was supported by a pile of money from a funding agency.}

### 2017-02-16 paper/paper.tex
s/McCliqueAlg/McSplitAlg/g

\begin{algorithm}[t]
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\medskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
    \nl    $M \gets M \cup \{(v,w)\}$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap \N(G, v) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap \N(H, w) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap \invN(G, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap \invN(H, w) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{Finding a maximum common subgraph.}
\label{McCliqueAlg}
\end{algorithm}

\paragraph{\cref{McCliqueAlg} in detail} The recursive procedure, $\FuncSty{expand}$, has
two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
each represented as a $\langle P, T \rangle$ pair as described above.  The
parameter $M$ is the current mapping of vertices.  On each call to
$\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
if and only if $v$ and $w$ belong to the same label class in $M$ ?? shouldn't that be future?.

@@

\begin{algorithm}[t]
\DontPrintSemicolon
\nl    \For {$l \in L$}{
\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} : u \neq v \wedge A_G[v][u] = l \}$ \;
\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} : u \neq w \wedge A_H[w][u] = l \}$ \;
\nl        \If {$P^{''} \negmedspace\neq \emptyset$ \bf{and} $T^{''} \negmedspace \neq \emptyset$}{
    \nl $\AlgVar{future^{'}} \gets \AlgVar{future^{'}} + \langle P^{''}\negmedspace, T^{''} \rangle$}
       }
    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McCliqueAlg} to handle directed and labelled cases.}
\label{labDirAlg}
\end{algorithm}

\paragraph{Undirected with edge labels} Each adjacency matrix entry contains an
edge label, or a null entry (non-adjacent) $0$.  We use \cref{labDirAlg}, by
letting $L$ be the union of $\{0\}$ with the set of all labels that appear in
the input graphs. Since there may be up to $g + h$ distinct labels, the loop in
\cref{labDirAlg} may execute up to $g + h$ times, resulting in $\BigO{(g+h)^2}$
time complexity per search node.  To achieve $\BigO{(g+h) \log (g+h)}$ time
complexity per search node, we can modify the algorithms to use sorting rather
than explicitly looping over all label classes, as follows.  First, run lines
17-19 of \cref{McCliqueAlg} to create a new label-class of vertices that are
not adjacent to $v$ or $w$, and remove these vertices from $\langle P',T'
\rangle$.  Next, sort $P'$ and $T'$ in ascending order of the label on the edge
from $v$ or $w$ to each vertex. We can then create the label classes
corresponding to each edge label by simultaneously traversing $P'$ and $T'$
from left to right.

@@

The basic \McSplit\ is beaten by the $k\downarrow$ algorithm of
\citet{UpcomingAAAIPaper} on this dataset. However, we can modify the \McSplit\
algorithm to use a top-down strategy similar to that used by $k\downarrow$  by
calling the main \FuncSty{McSplit} method once per goal size ($g, g-1, g-2,
\dots$); we backtrack (\lineref{PruneSearch} of \cref{McCliqueAlg}) when the
bound is strictly less than the goal size, and terminate when a solution of the
goal size is found. We expect that this could do well because in many cases the
matching covers nearly all of the smaller graph---indeed,
\cref{figure:sip-cumulative} shows that this approach is the strongest on these
instances.

@@

%%% This equivalence to CP-FC makes the correctness of \McSplit\ easy to establish:
%%% if we replace the label-class representation of $\mathit{future}$ with
%%% conventional domain stores, and replace \linerangeref{InnerLoop}{InnerLoopEnd}
%%% of \cref{McCliqueAlg} with \cref{cpAlg}.  The pruning of domains carried out in
%%% these lines can be easily seen to be equivalent to the corresponding lines in
%%% \cref{McCliqueAlg}.
%%% 
%%% 
%%% \begin{algorithm}
%%% \DontPrintSemicolon
%%% \nl    \For {$u \in \N(G, b)$}{
%%% \nl        remove $v$ from $D'_u$ \;
%%% \nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
%%%        }
%%% \nl    \For {$u \in \invN(G, b)$}{
%%% \nl        remove $v$ from $D'_u$ \;
%%% \nl        remove the neighbourhood of $t$ from $D'_u$ \;
%%%        }
%%% \caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm ?? Not sure about this bit, maybe we don't need it}
%%% \label{cpAlg}
%%% \end{algorithm}

>>>>

\begin{algorithm}[t]
\DontPrintSemicolon
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \;
\nl \Begin{
%\nl \lIf {$\AlgVar{future} = \emptyset$ \bf{and} $|M| > |\AlgVar{incumbent}|$}
\nl \lIf {$|M| > |\AlgVar{incumbent}|$}{$save(matching)$} \label{StoreIncumbent}
%\nl \lIf {$\AlgVar{future} = \emptyset$}{return}
\nl $bf \gets 0$ \label{SetBestFutureToZero} \;
\nl \lFor {$\langle P,T \rangle \in \AlgVar{future}$}{$\AlgVar{bf} \gets \AlgVar{bf} + \min(|P|,|T|)$}
\nl \lIf {$|M|  + \AlgVar{bf} \leq |\AlgVar{incumbent}|$}{return} \label{PruneSearch}
\medskip
\nl $\langle P,T \rangle \gets \FuncSty{SelectLabelClass}(\AlgVar{future})$ \label{SelectClass} \;
\nl $v \gets \FuncSty{SelectVertex}(P)$ \label{SelectVertex} \;
\nl \For {$w \in T$ \label{WLoop}} {
    \nl    $M \gets M \cup \{(v,w)\}$ \label{GrowM} \;
\nl    \bf{Let} $\AlgVar{future'} \gets \emptyset$ \label{NewFuture} \;
\nl    \For {$\langle P',T'\rangle \in future$ \label{InnerLoop}}{
\nl        \bf{Let} $P'' \gets P' \cap \N(G, v) \setminus \{v\}$ \label{NewPWithEdge} \;
\nl        \bf{Let} $T'' \gets T' \cap \N(H, w) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$ \label{AddToFutureWithEdge}}
\nl        \bf{Let} $P'' \gets P' \cap \invN(G, v) \setminus \{v\}$ \label{NewPWithoutEdge}  \;
\nl        \bf{Let} $T'' \gets T' \cap \invN(H, w) \setminus \{w\}$ \;
\nl        \If {$P'' \neq \emptyset$ \bf{and} $T'' \neq \emptyset$}{
\nl            $\AlgVar{future'} \gets \AlgVar{future'} + \langle P'' , T'' \rangle$} \label{InnerLoopEnd}
       }
\nl   $\FuncSty{expand}(\AlgVar{future'},M)$ \label{ExpandWithV} \;
\nl   $M \gets M \setminus \{(v,w)\}$ \label{ShrinkM} \;
  }
\nl $P' \gets P \setminus \{v\}$ \label{RemoveV} \;
\nl $\AlgVar{future} \gets \AlgVar{future} \setminus \{\langle P,T \rangle\}$\;
\nl \lIf {$P' \neq \emptyset$} {$\AlgVar{future} \gets \AlgVar{future} \cup \{\langle P',T \rangle \}$}
\nl $\FuncSty{expand}(\AlgVar{future},M)$ \label{ExpandWithoutV} \;
}
\;
\nl $\FuncSty{McSplit}(G,H)$ \label{McSplitFun} \;
\nl \Begin{
\nl $\AlgVar{incumbent} \gets \emptyset$ \;
\nl $\FuncSty{expand}(\{\langle V(G),V(H) \rangle \},\emptyset)$ \label{FirstExpandCall} \;
\nl return $|\AlgVar{incubent}|$ \;
}
\caption{Finding a maximum common subgraph.}
\label{McSplitAlg}
\end{algorithm}

\paragraph{\cref{McSplitAlg} in detail} The recursive procedure, $\FuncSty{expand}$, has
two parameters.  The parameter $\AlgVar{future}$ is a list of label classes,
each represented as a $\langle P, T \rangle$ pair as described above.  The
parameter $M$ is the current mapping of vertices.  On each call to
$\FuncSty{expand}$, the invariant holds that a $(v,w)$ pair may be added to $M$
if and only if $v$ and $w$ belong to the same label class in $M$ ?? shouldn't that be future?.

@@

\begin{algorithm}[t]
\DontPrintSemicolon
\nl    \For {$l \in L$}{
\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} : u \neq v \wedge A_G[v][u] = l \}$ \;
\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} : u \neq w \wedge A_H[w][u] = l \}$ \;
\nl        \If {$P^{''} \negmedspace\neq \emptyset$ \bf{and} $T^{''} \negmedspace \neq \emptyset$}{
    \nl $\AlgVar{future^{'}} \gets \AlgVar{future^{'}} + \langle P^{''}\negmedspace, T^{''} \rangle$}
       }
    \caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} of \cref{McSplitAlg} to handle directed and labelled cases.}
\label{labDirAlg}
\end{algorithm}

\paragraph{Undirected with edge labels} Each adjacency matrix entry contains an
edge label, or a null entry (non-adjacent) $0$.  We use \cref{labDirAlg}, by
letting $L$ be the union of $\{0\}$ with the set of all labels that appear in
the input graphs. Since there may be up to $g + h$ distinct labels, the loop in
\cref{labDirAlg} may execute up to $g + h$ times, resulting in $\BigO{(g+h)^2}$
time complexity per search node.  To achieve $\BigO{(g+h) \log (g+h)}$ time
complexity per search node, we can modify the algorithms to use sorting rather
than explicitly looping over all label classes, as follows.  First, run lines
17-19 of \cref{McSplitAlg} to create a new label-class of vertices that are
not adjacent to $v$ or $w$, and remove these vertices from $\langle P',T'
\rangle$.  Next, sort $P'$ and $T'$ in ascending order of the label on the edge
from $v$ or $w$ to each vertex. We can then create the label classes
corresponding to each edge label by simultaneously traversing $P'$ and $T'$
from left to right.

@@

The basic \McSplit\ is beaten by the $k\downarrow$ algorithm of
\citet{UpcomingAAAIPaper} on this dataset. However, we can modify the \McSplit\
algorithm to use a top-down strategy similar to that used by $k\downarrow$  by
calling the main \FuncSty{McSplit} method once per goal size ($g, g-1, g-2,
\dots$); we backtrack (\lineref{PruneSearch} of \cref{McSplitAlg}) when the
bound is strictly less than the goal size, and terminate when a solution of the
goal size is found. We expect that this could do well because in many cases the
matching covers nearly all of the smaller graph---indeed,
\cref{figure:sip-cumulative} shows that this approach is the strongest on these
instances.

@@

%%% This equivalence to CP-FC makes the correctness of \McSplit\ easy to establish:
%%% if we replace the label-class representation of $\mathit{future}$ with
%%% conventional domain stores, and replace \linerangeref{InnerLoop}{InnerLoopEnd}
%%% of \cref{McSplitAlg} with \cref{cpAlg}.  The pruning of domains carried out in
%%% these lines can be easily seen to be equivalent to the corresponding lines in
%%% \cref{McSplitAlg}.
%%% 
%%% 
%%% \begin{algorithm}
%%% \DontPrintSemicolon
%%% \nl    \For {$u \in \N(G, b)$}{
%%% \nl        remove $v$ from $D'_u$ \;
%%% \nl        remove the inverse neighbourhood of $t$ from $D'_u$ \;
%%%        }
%%% \nl    \For {$u \in \invN(G, b)$}{
%%% \nl        remove $v$ from $D'_u$ \;
%%% \nl        remove the neighbourhood of $t$ from $D'_u$ \;
%%%        }
%%% \caption{Replacement for \linerangeref{InnerLoop}{InnerLoopEnd} in CP algorithm ?? Not sure about this bit, maybe we don't need it}
%%% \label{cpAlg}
%%% \end{algorithm}
